Compararison of analytical and artificial neural network models in predicting hot deformation behavior of AISI P20 and interpolation-extrapolation capacity validation

### 2022-11-07 Article.tex
Update PTM model

The determination of the order $m$ of the polynomials defined by Equations (\ref{eq:ArA}-\ref{eq:Arn}) is defined according to the ability of the model to represent the non-linear dependence of the stress versus the strain and its generalization. 
The values $\ln\!A_i$, $\alpha_i$, $n_i$, and $Q_i$ ($i=1,2,3,\cdots,m$) are the coefficients of the polynomials to determine using a regression method. 
Setting $m=9$ gives the best results and the corresponding parameters are reported in Table \ref{tab:ARparameters}. 
\begin{table}[h!]
\centering
\caption{Parameters' values of the Arrhenius flow law for the P20 steel}
\begin{tabular}{llll}
\hline
$\alpha_i~(\text{MPa}^{-1})$ & $Q_i~(\text{MJ.mol}^{-1})$ & $\ln\!A_i~(\ps)$ & $n_i$\\
\hline
$\alpha_0=0.0392$ & $Q_0=0.467$ & $\ln\!A_0=35.8092$ & $n_0=4.8216$ \\
$\alpha_1=-0.4092$ & $Q_1=-0.6516$ & $\ln\!A_1=-58.8219$ & $n_1=3.2813$ \\
$\alpha_2=3.9229$ & $Q_2=7.6084$ & $\ln\!A_2=740.3302$ & $n_2=71.5963$ \\
$\alpha_3=-20.713$ & $Q_3=-48.016$ & $\ln\!A_3=-5049.2979$ & $n_3=-1956.1871$ \\
$\alpha_4=62.1376$ & $Q_4=66.795$ & $\ln\!A_4=11305.3573$ & $n_4=14460.9217$ \\
$\alpha_5=-99.9999$ & $Q_5=468.8898$ & $\ln\!A_5=20220.008$ & $n_5=-54309.8814$ \\
$\alpha_6=58.7108$ & $Q_6=-2303.1535$ & $\ln\!A_6=-153871.1135$ & $n_6=117612.0374$ \\
$\alpha_7=54.3688$ & $Q_7=4370.6566$ & $\ln\!A_7=317981.5396$ & $n_7=-148819.8996$ \\
$\alpha_8=-99.9999$ & $Q_8=-3939.3826$ & $\ln\!A_8=-297254.899$ & $n_8=102387.2537$ \\
$\alpha_9=42.5233$ & $Q_9=1397.0283$ & $\ln\!A_9=107585.17$ & $n_9=-29621.1293$ \\
\hline
\label{tab:ARparameters}
\end{tabular}
\end{table}
This modified form of the Arrhenius behavior law allows for accurate and reliable prediction over a wide range of temperatures and strain rates. 
Equation (\ref{eq:ArStress}) is finally used to compute the flow stress $\sigma^y$ from the strain $\varepsilon$, the strain rate $\mdot\varepsilon$ and the temperature $T$: 
\begin{equation}
\sigma^y = \frac{1}{\alpha(\varepsilon)} \ln\left\{\left(\frac{\mdot\varepsilon \exp{\left(\frac{Q(\varepsilon)}{RT}\right)}}{A(\varepsilon)}\right)^{\frac{1}{n(\varepsilon)}} + \left[1 + \left(\frac{\mdot\varepsilon \exp{\left(\frac{Q(\varepsilon)}{RT}\right)}}{A(\varepsilon)}\right)^{\frac{2}{n(\varepsilon)}}\right]^{\frac{1}{2}}\right\}
\label{eq:ArStress}
\end{equation}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpAR}
\caption{Comparison between the experimental and predicted flow stresses by AR model}
\label{fig:iCorrelationAR}
\end{figure}
\FloatBarrier
%-------------------------------------------------------------------------
\subsection{PTM model\label{sec:ProposedModel}}
%--------------------------------------------------------------------------
The PTM model is a generalized formulation of the MZA model. 
Indeed, the shortcomings of the MZA model are taken into account in this formulation, which makes the model flexible for any type of studied material as it does not require a limited number of parameters. The intrinsic functions of the model, which are the polynomial type functions, will determine the parameters of the model according to the degree of the polynomial, which provides a good fit for each function. The equation describing the PTM model is therefore given by.
\begin{equation}
\label{eq:PTM-model}
\sigma^y(\varepsilon^p,\mdot\varepsilon,T) = \left(\sum_{i=0}^{q}{A_i\varepsilon^{p^i}}\right) \exp\left[\left(\sum_{j=0}^{r}{B_j\varepsilon^{p^j}}\right)\left(T-T_0\right) + \left(\sum_{k=0}^{s}\left(\sum_{l=0}^{t}{C_k^l\varepsilon^{p^l}} \right)\left(T-T_0\right)^k \right)\ln\left( \frac{\mdot\varepsilon}{\mdot{\varepsilon}_0}\right)\right]
\end{equation}
where $A_i, B_j, C_k^l$ are the parameters (Table \ref{tab:PTM-parameters}) of the model to be identified using the procedure proposed in  \cite{TizeMha-2022}. Quantities $q$, $r$, $s$ and $t$ define the degree of the polynomials used to describe the behavior of the material. The larger these quantities are, the more parameters need to be identified for the PTM model.
The determination of the parameters of this model are calculated thanks to the LMFIT Python library \cite{Newville-2016} and for more details about this model  we invite the reader to read our previous work \cite{TizeMha-2022}. Thus all the parametrers of this model are calcuted with $q=5$, $r=5$, $s=1$ and $t=5$.

Comparison of predicted values of the PTM model and experimental values is shown in Figure \ref{fig:iCorrelationPTM}. The PTM model is suitable to describe the ﬂow behavior of AISI P20 steel, but for the strain rate $\mdot\varepsilon = 0.1\ \ps$ the prediction in not quite good. The deviation between the predicted values and the experimental values is relatively good.
\begin{table}[!h]
\centering{}
\caption{Parameters of PTM model\label{tab:PTM-parameters}}
\begin{tabular}{cc}
\hline
&   \\
Parameters &Values\\
&   \\
\hline
$A_0 (\text{MPa})$ & $62.314$ \\
$A_1 (\text{MPa})$ & $559.827$ \\
$A_2 (\text{MPa})$ & $-2160.03$ \\
$A_3 (\text{MPa})$ & $4641.94,$ \\
$A_4 (\text{MPa})$ & $-5193.06$ \\
$A_5 (\text{MPa})$ & $2379.1$ \\
$B_0 $   & $-0.0035805$\\
$B_1 $   & $0.00282334$ \\
$B_2 $   & $-0.00775301$\\
$B_3 $   & $0.0085135$ \\
$B_4 $   & $-0.00412815$\\
$C_0^0 $ & $0.148304$ \\
$C_0^1 $ & $-0.301958$ \\
$C_0^2 $ & $2.2706$ \\
$C_0^3 $ & $-5.33003$ \\
$C_0^4 $ & $5.48971$ \\
$C_0^5 $ & $-2.11566$\\
$C_1^0 $ & $-3.74842\ \times10^{-7}$\\
$C_1^1 $ & $0.00165233$\\
$C_1^2 $ & $-0.00270485$ \\
$C_1^3 $ & $-0.000483867$\\
$C_1^4 $ & $0.00174818$ \\
\hline
\end{tabular}
\end{table}
\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpPTM}
\caption{Comparison between the experimental and predicted flow stresses by PTM model }
\label{fig:iCorrelationPTM}
\end{figure}
\FloatBarrier

>>>>

The determination of the order $m$ of the polynomials defined by Equations (\ref{eq:ArA}-\ref{eq:Arn}) is defined according to the ability of the model to represent the non-linear dependence of the stress versus the strain and its generalization.
The values $\ln\!A_i$, $\alpha_i$, $n_i$, and $Q_i$ ($i=1,2,3,\cdots,m$) are the coefficients of the polynomials to determine using a regression method.
Setting $m=9$ gives the best results and the corresponding parameters are reported in Table \ref{tab:ARparameters}.
\begin{table}[h!]
\centering
\caption{Parameters' values of the Arrhenius flow law for the P20 steel}
\begin{tabular}{llll}
\hline
\multicolumn{1}{c}{$\alpha_i~(\text{MPa}^{-1})$} & \multicolumn{1}{c}{$Q_i~(\text{MJ.mol}^{-1})$} & \multicolumn{1}{c}{$\ln\!A_i~(\ps)$} & \multicolumn{1}{c}{$n_i$}\\
\hline
$\alpha_0=0.0392$ & $Q_0=0.467$ & $\ln\!A_0=35.8092$ & $n_0=4.8216$ \\
$\alpha_1=-0.4092$ & $Q_1=-0.6516$ & $\ln\!A_1=-58.8219$ & $n_1=3.2813$ \\
$\alpha_2=3.9229$ & $Q_2=7.6084$ & $\ln\!A_2=740.3302$ & $n_2=71.5963$ \\
$\alpha_3=-20.713$ & $Q_3=-48.016$ & $\ln\!A_3=-5049.2979$ & $n_3=-1956.1871$ \\
$\alpha_4=62.1376$ & $Q_4=66.795$ & $\ln\!A_4=11305.3573$ & $n_4=14460.9217$ \\
$\alpha_5=-99.9999$ & $Q_5=468.8898$ & $\ln\!A_5=20220.008$ & $n_5=-54309.8814$ \\
$\alpha_6=58.7108$ & $Q_6=-2303.1535$ & $\ln\!A_6=-153871.1135$ & $n_6=117612.0374$ \\
$\alpha_7=54.3688$ & $Q_7=4370.6566$ & $\ln\!A_7=317981.5396$ & $n_7=-148819.8996$ \\
$\alpha_8=-99.9999$ & $Q_8=-3939.3826$ & $\ln\!A_8=-297254.899$ & $n_8=102387.2537$ \\
$\alpha_9=42.5233$ & $Q_9=1397.0283$ & $\ln\!A_9=107585.17$ & $n_9=-29621.1293$ \\
\hline
\label{tab:ARparameters}
\end{tabular}
\end{table}
This modified form of the Arrhenius behavior law allows for accurate and reliable prediction over a wide range of temperatures and strain rates.
Equation (\ref{eq:ArStress}) is finally used to compute the flow stress $\sigma^y$ from the strain $\varepsilon$, the strain rate $\mdot\varepsilon$ and the temperature $T$:
\begin{equation}
\sigma^y = \frac{1}{\alpha(\varepsilon)} \ln\left\{\left(\frac{\mdot\varepsilon \exp{\left(\frac{Q(\varepsilon)}{RT}\right)}}{A(\varepsilon)}\right)^{\frac{1}{n(\varepsilon)}} + \left[1 + \left(\frac{\mdot\varepsilon \exp{\left(\frac{Q(\varepsilon)}{RT}\right)}}{A(\varepsilon)}\right)^{\frac{2}{n(\varepsilon)}}\right]^{\frac{1}{2}}\right\}
\label{eq:ArStress}
\end{equation}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpAR}
\caption{Comparison between the experimental and predicted flow stresses by AR model}
\label{fig:iCorrelationAR}
\end{figure}

%-------------------------------------------------------------------------
\subsection{PTM model\label{sec:PTM}}
%--------------------------------------------------------------------------

The PTM model \cite{TizeMha-2022} is a generalized formulation of the MZA model presented in Section \ref{sec:MZA}.
When establishing its formulation, the main shortcomings of the MZA model were taken into account, in order to make the PTM model flexible for any type of material studied because there is no need for a limited number of parameters as in the MZA model. Its construction is based on the use of polynomial functions as in the AR model.
Thus, the intrinsic functions of the model, of polynomial type, allow to adjust the model according to the degree selected for each of the 4 constituent polynomials, which provides a good fit for each function.
The equation describing the PTM model is therefore given by :
\begin{equation}
\label{eq:PTM-model}
\sigma^y = \left(\sum_{i=0}^{q}{A_i\varepsilon^{p^i}}\right) \exp\left[\left(\sum_{j=0}^{r}{B_j\varepsilon^{p^j}}\right)\left(T-T_0\right) + \left(\sum_{k=0}^{s}\left(\sum_{l=0}^{t}{C_k^l\varepsilon^{p^l}} \right)\left(T-T_0\right)^k \right)\ln\left( \frac{\mdot\varepsilon}{\mdot{\varepsilon}_0}\right)\right]
\end{equation}
where $A_i, B_j, C_k^l$ are the parameters (Table \ref{tab:PTM-parameters}) of the model to be identified using the procedure proposed in  \cite{TizeMha-2022}.
Quantities $q$, $r$, $s$ and $t$ define the degree of the polynomials used to describe the behavior of the material.
The larger these quantities are, the more parameters need to be identified for the PTM model.
The determination of the parameters of this model are calculated thanks to the LMFIT Python library \cite{Newville-2016} and for more details about this model  we invite the reader to read our previous work \cite{TizeMha-2022}.
Thus all the parametrers of this model are calcuted with $q=5$, $r=5$, $s=1$ and $t=5$.

Comparison of predicted values of the PTM model and experimental values is shown in Figure \ref{fig:iCorrelationPTM}.
The PTM model is suitable to describe the ﬂow behavior of AISI P20 steel, but for the strain rate $\mdot\varepsilon = 0.1\ \ps$ the prediction in not quite good.
The deviation between the predicted values and the experimental values is relatively good.

\begin{table}[h!]
\centering
\caption{Parameters' values of the PTM flow law for the P20 steel}
\begin{tabular}{llll}
\hline
\multicolumn{1}{c}{$A_i$} & \multicolumn{1}{c}{$B_i$} & \multicolumn{1}{c}{$C_0^i$} & \multicolumn{1}{c}{$C_1^i$}\\
\hline
$A_0=16.8528$ & $B_0=-0.0035$ & $C_0^0=0.1607$ & $C_1^0=0.0$\\
$A_1=340.6451$ & $B_1=-0.0131$ & $C_0^1=-0.6201$ & $C_1^1=0.0026$\\
$A_2=-1959.4041$ & $B_2=-0.0044$ & $C_0^2=4.9515$ & $C_1^2=-0.0035$\\
$A_3=4835.9698$ & $B_3=0.2217$ & $C_0^3=-13.1693$ & $C_1^3=-0.0222$\\
$A_4=-5517.6094$ & $B_4=-0.4988$ & $C_0^4=15.2499$ & $C_1^4=0.0608$\\
$A_5=2405.8399$ & $B_5=0.3211$ & $C_0^5=-6.5869$ & $C_1^5=-0.0412$\\
\hline
\label{tab:PTM-parameters}
\end{tabular}
\end{table}

\begin{figure}[!ht]
\centering
\includegraphics[width=1.02\columnwidth]
{Figures/CompExpPTM}
\caption{Comparison between the experimental and predicted flow stresses by PTM model }
\label{fig:iCorrelationPTM}
\end{figure}
\FloatBarrier
